PLAE enables quick and easy searching across one million ocular transcriptomes

### 2022-07-07 manuscript_00.Rmd
tweak fig sizes

```{r, fig.height=11, fig.width=6}
# tic()
# g <- consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% slice_max(n=8, order_by = `mean log2FC`) %>% mutate(g = paste0(Symbol, ' (', ID, ')')) %>% pull(g)
# 
# consist_diff_long <- scEiaD_2020_v01 %>%
#   tbl("diff_testing") %>%
#   filter(Gene %in% g) %>%
#   collect() %>%
#   filter(Base %in% ws_oct,
#          Group == 'CellType (Predict)',
#          Against == 'All') %>%
#   select(Gene, Organism, log2FoldChange, Base)
# 
# consist_diff_wide <- consist_diff_long %>% 
#   mutate(GO = paste0(Base, '_', Organism)) %>% 
#   select(-Organism, -Base) %>% 
#   pivot_wider(values_from=log2FoldChange, names_from = GO)
# consist_diff_celltype <- colnames(consist_diff_wide[,-1])
# consist_diff_wide <- consist_diff_wide %>% data.frame()
# row.names(consist_diff_wide) <- consist_diff_wide$Gene
# 
# consist_diff_wide <- consist_diff_wide[,-1]
# 
# 
# 
# consist_diff_organism <- str_extract(consist_diff_celltype, '_\\w+ \\w+') %>% gsub('_','',.)
# consist_diff_cts  <- str_extract(consist_diff_celltype, '.*_') %>% gsub('_','',.)
# consist_diff_genes <- row.names(consist_diff_wide) %>% gsub(' \\(.*','',.)
# 
# consist_diff_base <- consist_diff_genes %>% as_tibble() %>% rename(Symbol = value) %>% left_join(consist_diff %>% select(Symbol, Base) %>% unique()) %>% pull(Base)

consist_diff_hm <- ComplexHeatmap::Heatmap(consist_diff_wide, column_labels = consist_diff_organism, 
                              column_split = consist_diff_cts, 
                              row_split = consist_diff_base, show_row_dend = FALSE, 
                              row_labels = consist_diff_genes, 
                              column_title_rot = 90, 
                              name = 'log2FC', use_raster = TRUE,
                              row_title_rot = 0)
ComplexHeatmap::draw(consist_diff_hm, padding = unit(c(2, 2, 32, 2), "mm"), row_title = "Marker Genes")
#toc()
```

@@

Neurogenic and RPC have very similar expression patterns. We leverage our more specific differential test to directly assess genes were are different between neurogenic cells and RPC. We show the top 5 genes that are both more highly expressed in RPC relative to neurogenic cell and vice versa.
```{r, fig.width=9, fig.height=3}
# tic()
# pairwise_neurogenic_test<-  scEiaD_2020_v01 %>% 
#   tbl("diff_testing") %>% 
#   filter(Against  == "RPC",   
#          Group == 'CellType (Predict)', 
#          Base == "Neurogenic Cell") %>% 
#   group_by(Base, Against, Gene) %>% 
#   summarise(`mean log2FC` = mean(log2FoldChange), `mean padj` = mean(padj), `sig count` = sum(pvalue < 1e-2), `na count` = sum(is.na(log2FoldChange))) %>% 
#   arrange(-`mean log2FC`) %>%   
#   collect() %>% 
#   filter(!is.na(`mean log2FC`))
# 
# 
# # mutate(GENEID = str_extract(Gene, 'ENS\\w+')) %>% 
# # left_join(gene_tab, by = 'GENEID') %>% 
# # relocate(SYMBOL, GENEID) %>% 
# # dplyr::rename(Symbol = SYMBOL, ID = GENEID) %>% 
# # arrange(-`mean log2FC`)
# 
# down <- pairwise_neurogenic_test %>% filter(`sig count` > 1) %>% arrange(`mean padj`) %>% filter(`mean log2FC` < 0) %>% head(5) %>% pull(Gene)
# up <-  pairwise_neurogenic_test %>% filter(`sig count` > 1) %>% arrange(`mean padj`) %>% filter(`mean log2FC` > 0) %>% head(5) %>% pull(Gene)
source('src/make_exp_plot.R')

@@

input$exp_plot_genes <- up
exp_plot_iovs <- make_exp_plot(input, scEiaD_2020_v01, meta_filter)
a_legend <- get_legend(exp_plot_giga)
plot_grid(plot_grid(exp_plot_giga + theme(legend.position="none"),
          exp_plot_iovs + theme(legend.position="none") + ylab(''), labels = c('a','b')),
          a_legend, ncol = 1, rel_heights = c(1,0.2))
#toc()
```

```{r, fig.width=3, fig.height=7}
# tic()
# # flip_flop diff genes
# down_tab <- down_tab_big %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(Base %in% ws_oct, 
#          Against == 'All')
# 
# org_diff_diff_genes <- diff_tab_big %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(gene_ct_combo %in% down_tab$gene_ct_combo, 
#          Against == 'All') %>% 
#   arrange(padj) %>% mutate(x = paste0(Symbol, ' (', ID, ')')) 
# 
# flip_flop_genes <- scEiaD_2020_v01 %>% tbl('diff_testing_genes') %>% collect() %>% filter(grepl(org_diff_diff_genes %>% pull(ID) %>% paste0(., collapse = '|'), Gene)) %>% pull(Gene)
# flip_flop_long <- scEiaD_2020_v01 %>%
#   tbl("diff_testing") %>%
#   filter(Gene %in% genes) %>%
#   collect() %>%
#   mutate(Symbol = str_extract(Gene, '^\\w+ ') %>% gsub(' $', '', .)) %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(gene_ct_combo %in% org_diff_diff_genes$gene_ct_combo,
#          Group == 'CellType (Predict)',
#          Against == 'All') %>%
#   select(Gene, Organism, log2FoldChange, Base)
# 
# flip_flop_wide <- flip_flop_long %>% pivot_wider(values_from = log2FoldChange, names_from = Organism)
# flip_flop_organism <- colnames(flip_flop_wide)[c(3,4,5)]
# flip_flop_wide <- flip_flop_wide %>% data.frame()
# row.names(flip_flop_wide) <- paste0(flip_flop_wide$Gene,'_', flip_flop_wide$Base)
# 
# flip_flop_wide <- flip_flop_wide[,-c(1,2)]
# 
# 
# 
# flip_flop_genes <- row.names(flip_flop_wide) %>% gsub(' \\(.*','',.)
# 
# flip_flop_base <- row.names(flip_flop_wide) %>% gsub('.*_','',.)

>>>>

```{r, fig.height=24, fig.width=12}
# tic()
# g <- consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% slice_max(n=8, order_by = `mean log2FC`) %>% mutate(g = paste0(Symbol, ' (', ID, ')')) %>% pull(g)
# 
# consist_diff_long <- scEiaD_2020_v01 %>%
#   tbl("diff_testing") %>%
#   filter(Gene %in% g) %>%
#   collect() %>%
#   filter(Base %in% ws_oct,
#          Group == 'CellType (Predict)',
#          Against == 'All') %>%
#   select(Gene, Organism, log2FoldChange, Base)
# 
# consist_diff_wide <- consist_diff_long %>% 
#   mutate(GO = paste0(Base, '_', Organism)) %>% 
#   select(-Organism, -Base) %>% 
#   pivot_wider(values_from=log2FoldChange, names_from = GO)
# consist_diff_celltype <- colnames(consist_diff_wide[,-1])
# consist_diff_wide <- consist_diff_wide %>% data.frame()
# row.names(consist_diff_wide) <- consist_diff_wide$Gene
# 
# consist_diff_wide <- consist_diff_wide[,-1]
# 
# 
# 
# consist_diff_organism <- str_extract(consist_diff_celltype, '_\\w+ \\w+') %>% gsub('_','',.)
# consist_diff_cts  <- str_extract(consist_diff_celltype, '.*_') %>% gsub('_','',.)
# consist_diff_genes <- row.names(consist_diff_wide) %>% gsub(' \\(.*','',.)
# 
# consist_diff_base <- consist_diff_genes %>% as_tibble() %>% rename(Symbol = value) %>% left_join(consist_diff %>% select(Symbol, Base) %>% unique()) %>% pull(Base)

consist_diff_hm <- ComplexHeatmap::Heatmap(consist_diff_wide, column_labels = consist_diff_organism, 
                              column_split = consist_diff_cts, 
                              row_split = consist_diff_base, show_row_dend = FALSE, 
                              row_labels = consist_diff_genes, 
                              column_title_rot = 90, 
                              name = 'log2FC', use_raster = TRUE,
                              row_title_rot = 0)
ComplexHeatmap::draw(consist_diff_hm, padding = unit(c(0, 0, 1.5, 0), "in"), row_title = "Marker Genes")
#toc()
```

@@

Neurogenic and RPC have very similar expression patterns. We leverage our more specific differential test to directly assess genes were are different between neurogenic cells and RPC. We show the top 5 genes that are both more highly expressed in RPC relative to neurogenic cell and vice versa.

```{r, fig.width=15, fig.height=6, echo = F, message = F}
# tic()
# pairwise_neurogenic_test<-  scEiaD_2020_v01 %>% 
#   tbl("diff_testing") %>% 
#   filter(Against  == "RPC",   
#          Group == 'CellType (Predict)', 
#          Base == "Neurogenic Cell") %>% 
#   group_by(Base, Against, Gene) %>% 
#   summarise(`mean log2FC` = mean(log2FoldChange), `mean padj` = mean(padj), `sig count` = sum(pvalue < 1e-2), `na count` = sum(is.na(log2FoldChange))) %>% 
#   arrange(-`mean log2FC`) %>%   
#   collect() %>% 
#   filter(!is.na(`mean log2FC`))
# 
# 
# # mutate(GENEID = str_extract(Gene, 'ENS\\w+')) %>% 
# # left_join(gene_tab, by = 'GENEID') %>% 
# # relocate(SYMBOL, GENEID) %>% 
# # dplyr::rename(Symbol = SYMBOL, ID = GENEID) %>% 
# # arrange(-`mean log2FC`)
# 
# down <- pairwise_neurogenic_test %>% filter(`sig count` > 1) %>% arrange(`mean padj`) %>% filter(`mean log2FC` < 0) %>% head(5) %>% pull(Gene)
# up <-  pairwise_neurogenic_test %>% filter(`sig count` > 1) %>% arrange(`mean padj`) %>% filter(`mean log2FC` > 0) %>% head(5) %>% pull(Gene)
source('src/make_exp_plot.R')

@@

input$exp_plot_genes <- up
exp_plot_iovs <- make_exp_plot(input, scEiaD_2020_v01, meta_filter)
a_legend <- get_legend(exp_plot_giga)
plot_grid(plot_grid(exp_plot_giga + theme(legend.position="none") + xlab(''),
          exp_plot_iovs + theme(legend.position="none") + ylab('') + xlab(''), labels = c('a','b')),
          a_legend, ncol = 1, rel_heights = c(1,0.2))
#toc()
```

```{r, fig.width=5, fig.height=15, echo = F, message = F}
# tic()
# # flip_flop diff genes
# down_tab <- down_tab_big %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(Base %in% ws_oct, 
#          Against == 'All')
# 
# org_diff_diff_genes <- diff_tab_big %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(gene_ct_combo %in% down_tab$gene_ct_combo, 
#          Against == 'All') %>% 
#   arrange(padj) %>% mutate(x = paste0(Symbol, ' (', ID, ')')) 
# 
# flip_flop_genes <- scEiaD_2020_v01 %>% tbl('diff_testing_genes') %>% collect() %>% filter(grepl(org_diff_diff_genes %>% pull(ID) %>% paste0(., collapse = '|'), Gene)) %>% pull(Gene)
# flip_flop_long <- scEiaD_2020_v01 %>%
#   tbl("diff_testing") %>%
#   filter(Gene %in% genes) %>%
#   collect() %>%
#   mutate(Symbol = str_extract(Gene, '^\\w+ ') %>% gsub(' $', '', .)) %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(gene_ct_combo %in% org_diff_diff_genes$gene_ct_combo,
#          Group == 'CellType (Predict)',
#          Against == 'All') %>%
#   select(Gene, Organism, log2FoldChange, Base)
# 
# flip_flop_wide <- flip_flop_long %>% pivot_wider(values_from = log2FoldChange, names_from = Organism)
# flip_flop_organism <- colnames(flip_flop_wide)[c(3,4,5)]
# flip_flop_wide <- flip_flop_wide %>% data.frame()
# row.names(flip_flop_wide) <- paste0(flip_flop_wide$Gene,'_', flip_flop_wide$Base)
# 
# flip_flop_wide <- flip_flop_wide[,-c(1,2)]
# 
# 
# 
# flip_flop_genes <- row.names(flip_flop_wide) %>% gsub(' \\(.*','',.)
# 
# flip_flop_base <- row.names(flip_flop_wide) %>% gsub('.*_','',.)
